---
title: "Krig interpolation1"
author: "Hejaz Navaser"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
library(sp)
library(raster)
library(gstat)
library(fields)
library(ggplot2)
library(geoR)
require(MBA)

```


## loading the data

```{r}
load("C:/Users/Yasaman/Desktop/Spatial/datiAll.RData")
head(wolf)
summary(wolf)

```

```{r}
# Q1.Choose the variogram model (justify your choice)

# First of all, the data must be changed into spatial form. 
spwolf = SpatialPointsDataFrame(coords =cbind(wolf$x,wolf$y), data=wolf)

extent.wolf <- extent(spwolf)
gwolf <- expand.grid(x = seq(from = round(extent.wolf@xmin),
                     to = round(extent.wolf@xmax), length.out = 20),
                     y = seq(from = round(extent.wolf@ymin),
                     to = round(extent.wolf@ymax),length.out = 20))

coordinates(gwolf) <- ~x + y

grid <- gwolf
# storing spatial data in the grid
gridded(grid) <- TRUE 

plot(grid, main = paste("20x20 Interpolation Grid of Wolfcamp data"),col = "grey")

plot(spwolf, add = TRUE, pch = 15, cex = 0.5, col = "green") 


head(wolf)
summary(wolf)
```


```{r}
regpiezo<-lm(piezo~x+y, data= spwolf)
summary(regpiezo)

#at first we get our emperical variogram 

vv<-variogram(piezo~x+y,data = spwolf)
plot(vv,type="b")

```

As we can see in the plot, It is almost a little similar to our ideal mode in which the sill be visible and we observe a trend of covariance that depend only on distance.

```{r}
# Now, we want to try differnet fitting on our variogram.
# first we use OLS Regression to fit our variogram with different models

vv1<-gstat::fit.variogram(variogram(piezo~x+y,data = spwolf),vgm("Exp"), fit.method = 6)
plot(vv,vv1)


vv2<-gstat::fit.variogram(variogram(piezo~x+y,data = spwolf),vgm("Sph"), fit.method = 6)
plot(vv,vv2)


vv3 <- gstat::fit.variogram.reml(piezo~x+y,data = spwolf, model = vgm(psil= 400, "Sph",range = 30,nugget=500))
plot(vv,vv3)

vv4 <- fit.variogram(vv, vgm(c("Exp", "Mat", "Sph")))
plot(vv,vv4)

```


```{r}

# we Compare variogram models through kriging and cross validation
# kriging minimizes MSE, so the smaller MSE, the better the model is


krige1<-krige.cv(formula=piezo~x+y,spwolf,model = vv1, nmax=40, nfold=10)
krige2<-krige.cv(formula=piezo~x+y,spwolf,model = vv2, nmax=40, nfold=10)
krige3<-krige.cv(formula=piezo~x+y,spwolf,model = vv3, nmax=40, nfold=10)
krige4<-krige.cv(formula=piezo~x+y,spwolf,model = vv4, nmax=40, nfold=10)

bubble(krige1,"residual")
bubble(krige2,"residual")
bubble(krige3,"residual")
bubble(krige4,"residual")


```

RMSE (should be smaller)


```{r}

mean(krige1$residual^2) 
mean(krige2$residual^2)
mean(krige3$residual^2)
mean(krige4$residual^2)

```

As we can see, the Exponential is slightly better for Kriging interpolation


```{r}
# Q2. Run a spatial interpolation using kriging in the likelihood framework.
# here, we can use our best variogram model (Exponential) in likelihood framework

newkrige<-krige(formula=piezo~x+y, spwolf,model = vv1, newdata = grid)

summary(newkrige)
head(newkrige)

plot(newkrige, trend = '1st')

```

plotting of our kriging interpolation

```{r}

loc=coordinates(grid)
xyz=mba.surf(cbind(loc,newkrige$var1.pred),100,100,sp=T)$xyz.est
spplot(xyz)



spplot(newkrige["var1.pred"], main = "kriging Interpolation")

spplot(newkrige["var1.var"], main = "kriging standard deviation")

```

now, we want to assess our prediction with confidence intervals first of we should show the point view of this interpolation


```{r}

pointkrige<-data.frame(x=gwolf$x,y=gwolf$y,pred=newkrige@data$var1.pred,var=newkrige@data$var1.var, std=sqrt(newkrige@data$var1.var))

ggplot(pointkrige)+geom_point(aes(x=x,y=y, fill = pred,col=pred), size=1)+
  scale_fill_gradient("estimated value",low = "blue", high = "pink")+
  scale_color_gradient("estimated value",low = "blue", high = "pink")


pointkrige$low<- pointkrige$pred-1.95*pointkrige$std
pointkrige$up<- pointkrige$pred+1.95*pointkrige$std

pointkrige$low
pointkrige$up

#showing the number of bad prediction(as we can see, there is not any bad predicted point)

sum(pointkrige$low<0)

```


