---
title: "Rainfall-estimation_krig1"
author: "Hejaz Navaser"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

require(gstat)
require(sp)
require(sf)
require(interp)
require(raster)
require(ggplot2)
require(ggmap)
require(coda)
require(maptools)

```


## Loading the data

```{r}
load("C:/Users/39351/Documents/spatial homeworks/datiAll.RData")
load("C:/Users/39351/Documents/spatial homeworks/quota32.RData")

```

## Preparing the data

```{r}

# Removing the data of Puglia from the dataset since we are focing only on Basalicata and Calabria regions
datib <- dati[dati$Nome_Regione == "Basilicata", ]
datic <- dati[dati$Nome_Regione == "Calabria", ]
dati1 <- rbind(datib, datic)
dati1 <- dati1[, c(-1, -2)]         # removing the id columns for not offering any extra information



# The data of 1973
data1<-dati1[dati1$anno==1973,c(4, 5, 6, 8)]

# The data of 1975
data2<-dati1[dati1$anno==1975,c(4, 5, 6, 8)]

## Removing the id form the data of Basalica nad Calabria
Bas <- Bas[, -1]
Cal <- Cal[, -1]

## plotting the elevation in Basalicata and Calabria regions in 1973 
plot <- ggplot(data=data1, aes(x=XUTM, y=YUTM), title="The Elevation in 1973")
layer <- geom_point(data=data1, aes(col=Quota))

plot + layer + geom_point(data=Bas, aes(x=XUTM, y=YUTM), size=0.7, col="lightblue") + geom_point(data=Cal, aes(x=XUTM,y=YUTM), size=0.7, col="lightgray") + scale_color_gradient(low = "cyan", high="blue")+ggtitle("The Elevation in 1973")


## plotting the elevation in Basalicata na Calabria regions in 1975 
plot <- ggplot(data=data2, aes(x=XUTM, y=YUTM))
layer <- geom_point(data=data2, aes(col=Quota))

plot + layer + geom_point(data=Bas, aes(x=XUTM, y=YUTM), size=0.7, col="lightblue") + geom_point(data=Cal, aes(x=XUTM,y=YUTM), size=0.7, col="lightgray") + scale_color_gradient(low = "cyan", high="blue")+ggtitle("The Elevation in 1975")

```

```{r}
# in the quota32.RData object we have a fine grid for both regions *Basalica, Calabria)

quotacal<-Calabriaquota32[,-1]    # Calabria region
quotabas<-Basquota32[,-1]         # Basalica region


# harmonising the column names 
colnames(quotacal)<-colnames(quotabas)
quota<-data.frame(rbind(quotabas,quotacal)) # put both data into a single dataframe

summary(quota$Quota)

```

```{r}
## Plotting the stations of  Basalicata and Calabria in 1973
plot <- ggplot(data=data1, aes(x=XUTM, y=YUTM))
layer <- geom_point(data=data1, aes(col=Quota))

plot+layer+geom_point(data = Bas, aes(x=XUTM,y=YUTM),size=0.7,col="lightblue")+
  geom_point(data=Cal,aes(x=XUTM,y=YUTM),size=0.7,col="lightblue")+
  geom_point(data=quota,aes(x=XUTM,y=YUTM),size=0.3,col="pink")+
  scale_color_gradient(low="cyan", high="blue")+ggtitle("1973")

## Plotting the stations of  Basalicata and Calabria in 1975
plot <- ggplot(data=data2, aes(x=XUTM, y=YUTM))
layer <- geom_point(data=data2, aes(col=Quota))

plot+layer+geom_point(data = Bas, aes(x=XUTM,y=YUTM),size=0.7,col="lightblue")+
  geom_point(data=Cal,aes(x=XUTM,y=YUTM),size=0.7,col="lightblue")+
  geom_point(data=quota,aes(x=XUTM,y=YUTM),size=0.3,col="pink")+
  scale_color_gradient(low="cyan", high="blue")+ggtitle("1975")

```


```{r}

## Dividing the elevation to 4 categories from plain area to mountain area
quota.cl<-cut(quota$Quota,breaks=quantile(quota$Quota),include.lowest=T)
eticq=as.character(levels(quota.cl))
quota.cl<-cut(quota$Quota,breaks=quantile(quota$Quota),include.lowest=T,labels=c('Flat', 'Dulating', 'Hilly', 'mountains'))


## plotting the stations based on elevation categories in 1973
ggplot()+geom_point(data = Bas, aes(x=XUTM,y=YUTM),size=0.7,col="lightgray")+
  geom_point(data=Cal,aes(x=XUTM,y=YUTM),size=0.7,col="lightgray")+
  geom_point(data=quota,aes(x=XUTM,y=YUTM,fill=quota.cl,col=quota.cl))+
  geom_point(data=data1,aes(x=XUTM,y=YUTM),col="blue",shape = 20, size=2.5)+
  scale_color_manual("elevation", values=c(5:2))+
  scale_fill_manual("elevation", values=c(5:2))+
  theme_light()+ggtitle("1973")


## plotting the stations based on elevation categories in 1975
ggplot()+geom_point(data = Bas, aes(x=XUTM,y=YUTM),size=0.7,col="lightgray")+
  geom_point(data=Cal,aes(x=XUTM,y=YUTM),size=0.,col="lightgray")+
  geom_point(data=quota,aes(x=XUTM,y=YUTM,fill=quota.cl,col=quota.cl))+
  geom_point(data=data2,aes(x=XUTM,y=YUTM),col="black",shape = 20, size=2.5)+
  scale_color_manual("elevation", values=c(5:2))+
  scale_fill_manual("elevation", values=c(5:2))+
  theme_light()+ggtitle("1975")
```


```{r}

## to check if a linear relationship (trend) fits the data of 1973
attach(data1)
yy_1<-lm(totanno~XUTM+YUTM+Quota, data= data1)
summary(yy_1)


## to check if a linear trend fits the data of 1975
attach(data2)
yy_2<-lm(totanno~XUTM+YUTM+Quota, data=data2)
summary(yy_1)
```


```{r}

## plotting the result of fitting a linear trend to our data (diagnostic plots) 
plot(yy_1, main="1973")

plot(yy_2, main="1975")

```
```{r}
##plotting the observed surface 
xy<- interp(XUTM,YUTM,totanno, duplicate = 'mean')
fields::surface(xy)


lines(Cal[,1]/1000,Cal[,2]/1000,lwd=2,col="white")
lines(Bas[,1]/1000,Bas[,2]/1000,lwd=2,col = "white")

```

### Checking the distribution of the data
```{r}
## To garin a better results on our estimation we rescale the coordinates to km
data1$YUTM <- data1$YUTM / 1000
data1$XUTM <- data1$XUTM / 1000
data2$YUTM <- data2$YUTM / 1000
data2$XUTM <- data2$XUTM / 1000

#in order to be certain about the distribution of the rainfall amount, we perform the Shapiro-Wilk Normality test on the original rainfall data and the log-transformed data of 1973 and compare them

shapiro.test(data1$totanno)

transf<-MASS::boxcox(totanno~XUTM+YUTM+Quota,data = data1)
lambda<-transf$x[which(transf$y==max(transf$y))]    #very close to zero, so log transformation is the best choice 

#shapiro test on the transformed data
shapiro.test((data1$totanno^lambda-1)/lambda)
shapiro.test(log(data1$totanno)) # has the same result

```

The p-value obtained by the test shows that data are not normal, therefore, we performed the test on the log-transformed data, and as it is can be observed from the p-value the log data are normally distributed and can be used instead of the original data for estimation in Krging model. 



```{r}

# shapiro-wilk test on the original data 1975
shapiro.test(data2$totanno)#not Gaussian
transf<-MASS::boxcox(totanno~XUTM+YUTM+Quota,data = data2)
lambda<-transf$x[which(transf$y==max(transf$y))]#very close to zero that is very close to a log transformation

#shapiro test on the transformed data
shapiro.test((data2$totanno^lambda-1)/lambda)#Gaussian 


```


```{r}
# tranforming the data 1973
data1$totT<-(data1$totanno^lambda-1)/lambda


# tranforming the data 1975
data2$totT<-(data2$totanno^lambda-1)/lambda

```

```{r}
# using linear trend for the Empirical variogram of 1973
trend.data1<-formula(~XUTM+YUTM+Quota)

## Empirical variogram provide us with information of how the rainfall data is correlated with distance and elevation ( shows the covariance)
coordinates(data1) <- ~XUTM+YUTM
v_1 <- variogram(totT~YUTM+XUTM+Quota,data = data1)
plot(v_1,type="b", main="Variogram of 1973")

# Histogram of Residuals
hist(yy_1$residuals, main="REsidual of 1973", col='darkblue')
```




## Estimate the total amount of rain on the Calabria-Basilicata area using kriging in a maximum likelihood framework. (choose variogram and trend)

```{r}
# estimating some theoretical variogram for 1973

# using exponential distribution
vv1.e.ols <-gstat::fit.variogram(variogram(totT~YUTM+XUTM+Quota,data = data1),vgm("Exp"), fit.method = 6)
plot(v_1,vv1.e.ols, main="Variogram with Exponential - 1973")

# using spherical distribution 
vv1.s.ols<-gstat::fit.variogram(variogram(totT~YUTM+XUTM+Quota,data = data1),vgm("Sph"), fit.method = 6)
plot(v_1,vv1.s.ols, main="Variogram with  spherical - 1973")

vv1.s.reml <- gstat::fit.variogram.reml(totT~YUTM+XUTM+Quota,data = data1, model = vgm(psil= 400, "Sph",range = 30,nugget=500))

plot(v_1,vv1.s.reml)

```
#### Apperently, the Exponential variogram has the best fit on the data.

```{r}
# to compare the performance of each model and decide which one has the best performance on fitting variogram we used cross-validation 

# cross-validation on the Exponential model- 1973
k_exp_1<-krige.cv(formula=totT~YUTM+XUTM+Quota,data1,model = vv1.e.ols, nmax=40, nfold=10)


bubble(k_exp_1,"residual")
mean(k_exp_1$residual^2)

# cross- validation on the spherical model - 1973
k_sph_1<-krige.cv(formula=totT~YUTM+XUTM+Quota,data1,model = vv1.s.ols, nmax=40, nfold=10)

bubble(k_sph_1,"residual")
mean(k_sph_1$residual^2)
```

```{r}
# using linear trend for our Empirical variogram of 1975
trend.data2<-formula(~XUTM+YUTM+Quota)

## Empirical variogram provide us with information of how the rainfall data is correlated with distance and elevation ( shows the covariance)
coordinates(data2) <- ~XUTM+YUTM

v_2 <-variogram(totT~YUTM+XUTM+Quota,data = data2)
plot(v_2,type="b", main="Variogram of 1975")

# Histogram of residuals
hist(yy_2$residuals, main="REsidual of 1975", col='darkblue')

```

```{r}
# estimating some theoretical variogram for 1975

# using exponential distribution
vv2.e.ols <-gstat::fit.variogram(variogram(totT~YUTM+XUTM+Quota,data = data2),vgm("Exp"), fit.method = 6)
plot(v_2,vv2.e.ols, main="Variogram with Exponential - 1975")

# using spherical distribution 
vv2.s.ols<-gstat::fit.variogram(variogram(totT~YUTM+XUTM+Quota,data = data2),vgm("Sph"), fit.method = 6)
plot(v_2,vv1.s.ols, main="Variogram with spherical- 1975")
                            
vv2.s.reml <- gstat::fit.variogram.reml(totT~YUTM+XUTM+Quota,data = data2, model = vgm(psil= 400, "Sph",range = 30,nugget=500))
plot(v_2,vv2.s.reml)

```
##### for 1975 the exponential has the best fit on the data compared to the other variogram as well.

```{r}
# to compare the performance of each model model and decide which one has the best performance on fitting variogram we used cross-validation 

# cross-validation on the Exponential model- 1975
k_exp_2<-krige.cv(formula=totT~YUTM+XUTM+Quota,data2,model = vv2.e.ols, nmax=40, nfold=10)

bubble(k_exp_2,"residual")
mean(k_exp_2$residual^2)

# cross- validation on the spherical model - 1975
k_sph_2<-krige.cv(formula=totT~YUTM+XUTM+Quota,data2,model = vv2.s.ols, nmax=40, nfold=10)

bubble(k_sph_2,"residual",  na.rm=T)
mean(k_sph_2$residual^2)

```


```{r}
## we estimate the kriging surface now
quota.sp<-quota
quota.sp$XUTM<-quota$XUTM/1000
quota.sp$YUTM<-quota$YUTM/1000

coordinates(quota.sp) <- ~XUTM+YUTM
```

####  ML kriging For 1973
```{r}
ksurf1<-krige(formula=totanno~YUTM+XUTM+Quota,data1,model = vv1.e.ols,newdata = quota.sp)

ksurf.d1<-data.frame(XUTM=quota$XUTM/1000,YUTM=quota$YUTM/1000,pred=ksurf1@data$var1.pred,var=ksurf1@data$var1.var, std=sqrt(ksurf1@data$var1.var))

# building a %95 confidence intervals for the predictions of 1973
ksurf.d1$low<- ksurf.d1$pred-1.96*ksurf.d1$std
ksurf.d1$up<- ksurf.d1$pred+1.96*ksurf.d1$std


## to check if there are negative predictions among our estimations in 1973
sum(ksurf.d1$low < 0)
```

### mapping the results for 1973
```{r}

spplot(ksurf1["var1.pred"], main = "kriging interpolation", key.space=list(x=0.01,y=0.8,corner=c(-2,1)))

spplot(ksurf1["var1.var"], main = "kriging standard deviation", key.space=list(x=0.01,y=0.8,corner=c(-2,1)))


# plots with ggplot
ggplot(ksurf.d1)+geom_point(aes(x=XUTM,y=YUTM, fill = pred,col=pred), size=2.2)+
  geom_point(data = Bas, aes(x=XUTM/1000, y=YUTM/1000),col="lightblue")+
  geom_point(data = Cal, aes(x=XUTM/1000, y=YUTM/1000),col="lightblue")+
  theme_light()+
  scale_fill_gradient("Total rain 1973",low = "lightblue", high = "darkblue")+
  scale_color_gradient("Total rain 1973",low = "lightblue", high = "darkblue")
```

####  ML kriging For 1975
```{r}
ksurf2<-krige(formula=totanno~YUTM+XUTM+Quota,data2,model = vv2.e.ols,newdata = quota.sp)

ksurf.d2<-data.frame(XUTM=quota$XUTM/1000,YUTM=quota$YUTM/1000,pred=ksurf2@data$var1.pred,var=ksurf2@data$var1.var, std=sqrt(ksurf2@data$var1.var))

# building a %95 confidence intervals for the predictions of 1975 
ksurf.d2$low<- ksurf.d2$pred-1.96*ksurf.d2$std
ksurf.d2$up<- ksurf.d2$pred+1.96*ksurf.d2$std

## to check if there are negative predictions among our estimations in 1975
sum(ksurf.d2$low < 0)
```

### mapping the results for 1975
```{r}
spplot(ksurf2["var1.pred"], main = "kriging interpolation", key.space=list(x=0.01,y=0.8,corner=c(-2,1)))

spplot(ksurf2["var1.var"], main = "kriging standard deviation", key.space=list(x=0.01,y=0.8,corner=c(-2,1)))


# plots with ggplot
ggplot(ksurf.d2)+geom_point(aes(x=XUTM,y=YUTM, fill = pred,col=pred), size=2.2)+
  geom_point(data = Bas, aes(x=XUTM/1000, y=YUTM/1000),col="lightblue")+
  geom_point(data = Cal, aes(x=XUTM/1000, y=YUTM/1000),col="lightblue")+
  theme_light()+
  scale_fill_gradient("Total rain 1975",low = "lightblue", high = "darkblue")+
  scale_color_gradient("Total rain 1975",low = "lightblue", high = "darkblue")
```






