---
title: "rainfall analysis- data interpolation"
author: "Hejaz Navaser"
date:
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
require(gstat)
require(sp)
require(raster)
require(ggplot2)
require(MBA)
require(ggmap)
require(coda)
require(ggpubr)
require(psych)
require(vtable)

```

##Loading the data

```{r }

load("C:/Users/39351/documents/spatial homeworks/datiAll.RData")
load("C:/Users/39351/documents/spatial homeworks/quota32.RData")

```



## Creating the dataset for each year, the dataset is limited to two regions: Basalicata and Calabria

### Dataset1

```{r}
# Removing the data of Puglia from the dataset since we are focing only on Basalicata and Calabria regions
datib <- dati[dati$Nome_Regione == "Basilicata", ]
datic <- dati[dati$Nome_Regione == "Calabria", ]
dati1 <- rbind(datib, datic)

# The Dataset for 1973
data1 <- dati1[dati1$anno == 1973, ]
data1 <- data1[, c(-1, -2)]         # removing the id columns for not offering any extra information


# Taking the latitude, longitude, the elevation, and the amount of rain 
data_1 <- data1[,c(2, 3, 6, 8)]

# Coordinates based on latitude and longitude
coordinates(data_1) <- ~Lon10+Lat10+Quota

dataset1 <- data1
coordinates(dataset1) <- ~XUTM+YUTM+Quota

#region_info <- summary(data1$Nome_Regione)
#elevation_info <- summary(data1$Quota)
#rainfall_info <- summary(data1$totanno)

st(data1, vars=c('Nome_Regione','Quota','totanno'))

```

### Dataset2

```{r}
# The dataset for 1975 
data2 <- dati1[dati1$anno == 1975,]
data2 <- data2[, c(-1, -2)]     # removing the id columns 

data_2 <- data2[,c(2, 3, 6, 8)]

coordinates(data_2) <- ~Lon10+Lat10+Quota

dataset2 <- data2
coordinates(dataset2) <- ~XUTM+YUTM+Quota

#summary(data2$Nome_Regione)
#summary(data2$Quota)
#summary(data2$totanno)
st(data2, vars=c('Nome_Regione','Quota','totanno'))

```


## Exploratory Analysis

```{r}
## summary information on the data of 1973 based on each region
describeBy(data1, group=data1$Nome_Regione)

```

```{r}

## detect the outliers
lower_bound1 <- quantile(data1$totanno, 0.01) #set a lower bound
upper_bound1 <- quantile(data1$totanno, 0.99) #set an upper bound

outlier_ind1 <- which(data1$totanno < lower_bound1 | data1$totanno > upper_bound1)
outlier_ind1

data1[outlier_ind1,"totanno"] ## print the outliers

# boxplot of the rainfall 1975
boxplot(data1$totanno, col="lightskyblue", main= "Rainfall in 1973")


```

```{r}

# The Elevation distribution Histogram of 1973
{hist(data1$Quota, prob=TRUE, breaks=20, main="Elevation Distribution histogram of 1973", xlab= "Elevation",  col = "#0c4c8a") #plotting the elevation of 1973 
  
curve(dnorm(x, mean=mean(data1$Quota), sd=sd(data1$Quota)), add=TRUE) # draw the distribution line of the Elevation

abline(v=mean(data1$Quota), col="blue") # plotting the mean of the distribution
text(mean(data1$Quota),0.80,"Mean", col = "blue", adj = c(0, 1))

abline(v=median(data1$Quota), col="red") # plotting the median 
text(median(data1$Quota),0.80,"Median", col = "red", adj = c(0, 1))}


## The Rainfall Distribution Histogram
{hist(data1$totanno, prob=TRUE,breaks=20,main="Rainfall distribution of 1973", xlab= "Rainfall", col = "#0c4c8a") #plotting the rainfall of 1973
curve(dnorm(x, mean=mean(data1$totanno), sd=sd(data1$totanno)), add=TRUE) # draw the distribution line of the rainfall
abline(v=mean(data1$totanno), col="blue") # The distribution mean
text(mean(data1$totanno),0.80,"Mean", col = "blue", adj = c(0, 1)) 
abline(v=median(data1$totanno), col="red") # The distribution median
text(median(data1$totanno),0.80,"Median", col = "red", adj = c(0, 1))}

## perform the Shapiro-wilk test for normality
shapiro.test(data1$totanno)

```


As it can be seen the mean and the median are very close but we cannot say that the data is normally distributed, so we performed a **Shapiro-Wilk test** and the **p-value** is very small so as it is expected the data is not normally distributed. So, we are going to log-transform the data and perform the **shapiro test** on transformed data. 



```{r}

# perform Shapiro-Wilk Normality test on the log-transformed rainfall data of 1973
shapiro.test(log(data1$totanno)) #1973

# The Histogram of the log_transformed rainfall data
{hist(log(data1$totanno), prob=TRUE,breaks=20,main="Rainfall log-distribution of 1973", xlab= "Rainfall", col="purple")
curve(dnorm(x, mean=mean(log(data1$totanno)), sd=sd(log(data1$totanno))), add=TRUE) # plotting the distribution line of the rainfall
abline(v=mean(log(data1$totanno)), col="blue") # The distribution mean
text(mean(log(data1$totanno)),0.80,"Mean", col = "blue", adj = c(0, 1)) 
abline(v=median(log(data1$totanno)), col="red") # The distribution median
text(median(log(data1$totanno)),0.80,"Median", col = "red", adj = c(0, 1))}

```

as it is obvious from the histogram and the result of the **Shapiro-wilk test**, the transformed data are normally distributed



```{r}
## summary information on the data of 1973 based on each region
describeBy(data2, group=data2$Nome_Regione,digits = 2)

```

```{r boxplot 1975}
## detect the outliers
lower_bound2 <- quantile(data2$totanno, 0.01) #set a lower bound
upper_bound2 <- quantile(data2$totanno, 0.99) #set an upper bound

outlier_ind2 <- which(data2$totanno < lower_bound2 | data2$totanno > upper_bound2)
outlier_ind2

data2[outlier_ind2,"totanno"] ## print the outliers

# box-plot of the rainfall 1975
boxplot(data2$totanno, col="lightskyblue", main= "Rainfall in 1975")

```

```{r}
# The Elevation distribution Histogram of 1975
{hist(data2$Quota, prob=TRUE, breaks=20, main="Elevation Distribution histogram", xlab= "Elevation", col = "#0c4c8a") #plotting the elevation of 1975 
  
curve(dnorm(x, mean=mean(data2$Quota), sd=sd(data2$Quota)), add=TRUE) # draw the distribution line of the Elevation

abline(v=mean(data2$Quota), col="blue") # plotting the mean of the distribution
text(mean(data2$Quota),0.80,"Mean", col = "blue", adj = c(0, 1))

abline(v=median(data2$Quota), col="red") # plotting the median 
text(median(data2$Quota),0.80,"Median", col = "red", adj = c(0, 1))}


## The Rainfall Distribution Histogram of 1975
{hist(data2$totanno, prob=TRUE,breaks=20,main="Rainfall distribution", xlab= "Rainfall", col = "#0c4c8a") #plotting the rainfall of 1975
curve(dnorm(x, mean=mean(data2$totanno), sd=sd(data2$totanno)), add=TRUE) # draw the distribution line of the rainfall
abline(v=mean(data2$totanno), col="blue") # The distribution mean
text(mean(data2$totanno),0.80,"Mean", col = "blue", adj = c(0, 1)) 
abline(v=median(data2$totanno), col="red") # The distribution median
text(median(data2$totanno),0.80,"Median", col = "red", adj = c(0, 1))}

## perform Shapiro-test
shapiro.test(data2$totanno)

```

looking at the histogram and the results of **Shapiro test** on data of 1975, we can assume that the data are normally distributed and we will not use log-transform data, but for a comparison purpose we are going to plot the transformed data and perform a **shapiro test** on it.  



```{r}

# Perform Shapiro-Wilk Normality test on the log-transformed rainfall data of 1975
shapiro.test(log(data2$totanno)) #1975


# The Histogram of the log_transformed rainfall data
{hist(log(data1$totanno), prob=TRUE,breaks=20,main="Rainfall log-distribution of 1973", xlab= "Rainfall", col='purple')
curve(dnorm(x, mean=mean(log(data1$totanno)), sd=sd(log(data1$totanno))), add=TRUE) # plotting the distribution line of the rainfall
abline(v=mean(log(data1$totanno)), col="blue") # The distribution mean
text(mean(log(data1$totanno)),0.80,"Mean", col = "blue", adj = c(0, 1)) 
abline(v=median(log(data1$totanno)), col="red") # The distribution median
text(median(log(data1$totanno)),0.80,"Median", col = "red", adj = c(0, 1))} 

```



## Choose the optimal value for p in the IDW estimator using cross-validation

```{r finding best p value}

best_p <- function(dataset, p){
  # The goal : find the best power value p for IDW estimation LOOCV
  
  # Preparing the Dataset to fit IDW function
  if (typeof(dataset) == "list"){
  
    # converting the dataframe to SpatialPointsDataFrame
    
    dataset <- SpatialPointsDataFrame(data=dataset, coords=cbind(dataset$XUTM, dataset$YUTM, dataset$Quota)) 
    
  }
  
    # creation of the grid for the estimations
  grd <- as.data.frame(dataset)[,c(4,5,6)]
  coordinates(grd) <- ~ XUTM + YUTM + Quota
  
  
  # Applying LOOCV
  m <- 100000
  dd <- matrix(data = NA, nrow = length(p), ncol = 2)
  dd[,1] <- p
  
  
  # looping over the different powers (p values)
  for (j in p){ 
    
    
   RMSE <- c()
    
    # testing all the observations  of the train-set
    for (i in 1:length(dataset)){ 
      
      # We estimate the missing parameter
      IDW <- gstat::idw(totanno~1, dataset[-i,], dataset, idp=j) 
      
      # Save the RMSE (to keep the same dimension of the data)
      RMSE[j] <- sqrt(mean((dataset$totanno[j] - IDW$var1.pred)^2)) 

    }
    
    # Getting the mean of the results
    dd[j,2] <- RMSE[j]
  }
  bp <- dd[which.min(dd[,2]),1]
  return(list (dd, bp))
}

```


```{r best p value for 1973}
# finding best P value for 1973
k <- best_p(dataset1, p = seq(1,4,0.4))
k

```


```{r best p value for 1975 }

# finding best P value for 1975
k <- best_p(dataset2, p = seq(1,4,0.4))
k 

```


## Using the IDW estimator, map rainfall values on the grid recorded in the database quota

```{r}
# mapping the rainfall values for 1973
#creating the grid
colnames(Basquota32) <- colnames(Calabriaquota32)
grid <- rbind(Basquota32, Calabriaquota32)
?idw

# Plotting the coordinates of stations in Basalicata and Calibria  
plot(grid$X, grid$Y, pch = 20, col=8)
points(Bas$XUTM, Bas$YUTM,)
points(Cal$XUTM, Cal$YUTM)
points(data1$XUTM, data1$YUTM, col = 2, pch = 17) #1973

# converting the grid into a sp object
coordinates(grid) <- ~ X + Y + quota

# Using IDW to interpolate points using the best P value of 1975
new_points <- idw(totanno ~ 1, locations=dataset1, newdata=grid, idp = 1.4)
new_points.output <- as.data.frame(new_points)
names(new_points.output)[1:4] <- c("x", "y", "quota", "var1.pred")

gridd <- cbind(as.data.frame(grid)[2:3], new_points.output[4])


plot1 <- ggplot(data=as.data.frame(gridd),aes(x=x,y=y))
layer1 <- geom_point(data=new_points.output[,-5],aes(fill=var1.pred, col = var1.pred))

plot1+layer1+coord_equal()


```

```{r mapping rainfall values on grid 1975, , warning=FALSE}
# mapping the rainfall values of 1975 on grid 

## creating the grid
colnames(Basquota32) <- colnames(Calabriaquota32)
grid2 <- rbind(Basquota32, Calabriaquota32)


# Plotting the coordinates of stations in Basalicata and Calibria region 
plot(grid$X, grid$Y, pch = 20, col=8)
points(Bas$XUTM, Bas$YUTM,)
points(Cal$XUTM, Cal$YUTM)
points(data2$XUTM, data2$YUTM, col = 2, pch = 17) #1975

# converting the grid into a sp object
coordinates(grid2) <- ~ X + Y + quota

# Using IDW to interpolate points using the best P value
new_points2 <- idw(totanno ~ 1, locations=dataset2, newdata=grid2, idp = 1.8)
new_points2.output <- as.data.frame(new_points2)
names(new_points2.output)[1:4] <- c("x", "y", "quota", "var1.pred")
#new_points.output

gridd2 <- cbind(as.data.frame(grid)[2:3], new_points2.output[4])
#gridd

plot2 <- ggplot(data=as.data.frame(gridd2),aes(x=x,y=y))
layer2 <- geom_point(data=new_points2.output[,-5],aes(fill=var1.pred, col = var1.pred))

plot2+layer2+coord_equal()


```









